---
title: "Ingestion-Capstone"
output: html_document
date: "2026-02-04"
---

```{r}
setwd("C:/Users/wjcor/Downloads/Data_205")
```


```{r}
library(tidyverse)
library(reshape2)
```
```{r}
house=read.csv("C:/Users/wjcor/Downloads/Housing_Code_Violations_20260204.csv")
tax=read.csv("C:/Users/wjcor/Downloads/Real_Property_Tax_-_2024_20260204.csv")
```

The latitude and longitude columns for both datasets are empty and thus so is the "location" column. We remove those to keep clutter at a minimum

```{r}
house$Latitude=house$Latitude=NULL
house$Longitude= house$Longitude=NULL
house$Location=house$Location=NULL
house$address=house$Street.Address
```

Residence means whether the owner lives there, not important to our questions. We rename the address column to an easier name. 

```{r}
tax$RESIDENCE=tax$RESIDENCE=NULL
tax$address=tax$PROPERTY.ADDRESS
```

This joins the two datasets by their shared address variables

```{r}
joined=left_join(house, tax, by="address")
```
This removes NAs in the total column, which make up about a third of the data, we won't be removing other NA values, but these columns are largely useless for us. 


```{r}
joined_na=joined[is.na(joined$TOTAL)==FALSE,]
```

This  sets the characters we want to remove from our numerical data

```{r}
removechar="[$,]"
```

This removes the characters and sets these columns to numeric 

```{r}
fake_joined=joined_na
for (col in 1:ncol(fake_joined)){
  if(grepl("\\$", fake_joined[1,col])==TRUE){
    fake_joined[,col]=as.numeric(gsub(removechar, '', fake_joined[,col]))
  }
  }
```

Now we are creating a dataframe for tallying the violations from a table that automatically counts the frequency of each unique address

```{r}
housetable=data.frame(table(fake_joined$address))
```

This makes it easier to write the ensuing code by simplifying the names of columns 

```{r}
colnames(housetable)=c("address", "Violation_Count")
```

This sets our count variable to numeric 

```{r}
housetable$Violation_Count=as.numeric(housetable$Violation_Count)

```
This joins our large dataset with the violation count dataset

```{r}
fake_joined=left_join(x=fake_joined,y=housetable)
```

This removes a particularly egregious address, which, either by data entry mistake or sheer awfulness, has ~250,000 individual violations. 

```{r}
fake_joined=fake_joined|>
  filter(address !="14120 WEEPING WILLOW DR")
```

Now we are taking only the unqiue combinations of these variables, which should reduce the number of duplicate addresses. 

```{r}
uniquerows= fake_joined |>
  distinct(address, TAX.CLASS, TOTAL, CITY,Zip.Code,STATE,Violation_Count)
```

Now we are cutting the dataset into 3 distinct partitions and only the relevant variables. We will import these columns into a geocoding tool from the census bureau. This will result in latitude and longitude variables we can use for mapping.   

```{r}
cut1=uniquerows[1:9999,c(1,4,6,5)]
cut2=uniquerows[10000:19998,c(1,4,6,5)]
cut3=uniquerows[19999:22078,c(1,4,6,5)]
```

This exports the datasets as csvs

```{r}
write.csv(cut1,file="cut1.csv")
write.csv(cut2,file="cut2.csv")
write.csv(cut3,file="cut3.csv")
```

This takes our slightly modified results and imports them back into R

```{r}
part1=read.csv("C:/Users/wjcor/Downloads/GeocodeResults (1)(Table1_1).csv")
part2=read.csv("C:/Users/wjcor/Downloads/GeocodeResults (2)(Table1).csv")
part3=read.csv("C:/Users/wjcor/Downloads/GeocodeResults (3)(Table1).csv")
```

Now we change the column names in our dataset for easiest use


```{r}
colnames(part1)=c("Address","Longitude","Latitude")
```

And then connect the parts 

```{r}
connected=rbind(part1,part2,part3)
```

And then rename the connected dataset

```{r}
colnames(connected)=c("address","Longitude","Latitude")
```

This joins the unique values we seperated before with the latitude and longitude values

```{r}
small=left_join(x=uniquerows,y=connected)
```

This writes the finalized dataset into a csv for importing into Tableau

```{r}
write.csv(small, file="Small.csv")
```

This is a partition of the big dataset that takes all of the individual relevant columns from the big dataset for analysis on violation items and then renames them for ease of use. 


```{r}
item=data.frame("Item"=fake_joined$Item,"Tax Class"=fake_joined$TAX.CLASS, "zip"=fake_joined$Zip.Code,"Start Date"=fake_joined$Date.Filed,"End Date"=fake_joined$Date.Closed,"address"=fake_joined$Street.Address)

```

Finally, we write this to a csv so we can export. 

```{r}
write.csv(item,file="Item.csv")
```


